Revisiting the hydrodynamic modulation of short surface waves by longer waves

### 2024-10-01 paper/paper.tex
A few more references + spelling

All modulation solutions mentioned thus far are steady in the reference frame
of the long wave---short waves become shorter and higher (and thus, steeper)
preferentially on the long-wave crests.
Conversely, they elongate and flatten preferentially in the long-wave troughs.
Recently, \citet{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are appropriate by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario in
which a long-wave train suddenly appears to modulate a uniform field of short
waves.
The numerical simulations stabilize if the short-wave field is initialized
from existing steady solutions, like those by \citet{longuet1960changes}.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.
Looking at the conservation equations alone, it is not obvious that it is.

In this paper, we revisit this problem from linear wave theory but consider
the long-wave slope to be significant enough to require evaluating the
short-waves at the long-wave surface rather than at the mean water level.
This allows us to derive alternative, steady, non-linear solutions for the
modulation of short waves.
The analytical solutions yield modulations that are stronger than the original
solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
long-wave steepness up to $\approx 0.2$.
To validate the steadiness of the analytical solutions, we perform numerical
simulations of the full wave crest and action conservation laws, and compare
the results to the analytical solutions presented here, as well as the prior
numerical results.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarized in Table
\ref{table:modulation_literature_summary}.
We first present the governing equations in \S\ref{section:governing_equations},
and then the analytical solutions for the modulation of short waves by long waves
in \S\ref{section:analytical_solutions}.
Then, we describe the numerical simulations and their results in
\S\ref{section:numerical_solutions}.
Finally, we discuss the results and their implications in
\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.

@@

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
angular frequency.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearizing the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta$ and the
linearized conservation of wave crests in (\ref{eq:wavenumber}) results in
excess short-wave wavenumber being created.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.

@@

Let us now examine the contributions of the different terms to the short-wave
effective gravity modulation.
Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
a variety of the gravity modulation forms.
This value of $\varepsilon_L$ allows easier visualization of the modulation
differences, while still being relatively moderate.
To the first order in $\varepsilon_L$, the gravity modulation is out of phase
with the long-wave elevation, with the short waves experiencing a lower effective
gravity on the long-wave crests and a higher effective gravity in the troughs.
Evaluating the orbital accelerations at $z=\eta_L$ causes a gravity reduction
around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
the exponential change in the orbital accelerations away from the mean water level
(Fig. \ref{fig:effective_gravities}b).
The difference between the linear and Stokes long waves at this steepness is
negligible.
The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
Naturally, there is no curvature exactly at the crests and the troughs, and its
effects are largest at the front and rear faces of the long wave, where the
covariance between the local slope and the horizontal accelerations are largest.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalized by the modulation of the linear wave evaluated at $z=0$, $g_1$.
}
\label{fig:effective_gravities}
\end{figure}

@@

The modulation of short-wave amplitude can be derived in a similar way as we did
for the wavenumber, except that here we linearize the wave action balance
(\ref{eq:wave_action2}) and integrate it in time:

@@

To quantify the contribution of the nonlinear terms in (\ref{eq:wavenumber})
and (\ref{eq:wave_action2}), and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretized using
second-order central finite difference in space and integrated in time using the
fourth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, thus neglecting the Stokes drift
induced by long waves on short-wave groups
\citep{stokes1847,van2018stokes,monismith2020stokes}.
Their gravity modulation also does not consider the nonlinear or the curvature
effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the more sophisticated MUSCL4 scheme \citep{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The maximum relative error of the centered finite difference in this model
is $\approx 0.066\%$.
We show in the next section that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

Finally, the numerical equations here are integrated in a fixed reference frame,
rather than that moving with the long-wave phase speed.
Either approach produces equivalent modulation results, however a fixed
reference frame may allow for a more intuitive interpretation of the results.
This difference is important to keep in mind when comparing the results
of \citet{peureux2021unsteady} and those presented here; in their Fig. 1 the
long-wave phase is fixed and the short waves are moving leftward
with the speed of $C_{pL} - C_g - U$ (neglecting group speed inhomogeneity),
whereas in the figures that follow, the long wave is moving rightward with its
phase speed $C_{pL}$ and the short waves (that is, their action) are moving
rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).

@@

The first set of simulations is performed for the long-wave steepness of
$\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
The long-waves are initialized from rest ($a_L = 0$) and gradually ramped up
to their target steepness over 5 long-wave periods, thus emulating the passage
of a long-wave group
(the importance of which we discuss in more detail in the next subsection).
At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
and steepness modulation are all remarkably similar to the analytical solutions
(Fig. \ref{fig:numerical_solutions}a,d,g).
This suggests that at low long-wave steepness that the steady analytical solution
is a reasonable proxy for the fully-nonlinear crest-action conservation
equation solutions.
The numerical solutions diverge more notably from the analytical ones at
moderate steepness of 0.2 (Fig. \ref{fig:numerical_solutions}b,e,h), and are
considerably different at very high steepness of 0.4
(Fig. \ref{fig:numerical_solutions}c,f,i).
These differences suggest that at high steepness the inhomogeneity of
gravitational acceleration and group speed, otherwise ignored in the steady
solutions, become important.

@@

The key result of \citet{peureux2021unsteady} is that of unsteady steepening of
short waves when a homogeneous field of short waves is forced by an infinite
long-wave train.
The steepening is driven by the progressive increase in the action of short
waves (as opposed to their wavenumber), and is centered on the short-wave group
that begins its journey in the trough of the long wave.
This is because the unmodulated short waves that begin their journey in the
trough first enter the convergence region before reaching the divergence region.
Inversely, the short waves that begin around the crest of the long-wave first
enter the divergence region and will thus have their action and wavenumber
decreased.
We reproduce this behavior in Fig. \ref{fig:modulation_3panel_infinite}, which
shows the evolution of the short-wave action, wavenumber, and steepness
modulation (here defined as the change of a quantity relative to its initial value).
The long waves have $\varepsilon_L = 0.1$ and $k_L = 1$, and the short waves
have $k = 10$.
After 10 long-wave periods, the short-wave action is approximately doubled,
consistent with \citet{peureux2021unsteady}.
Another important property of this solution is that the short-wave group that
experiences peak amplification (dampening) are not locked-in to the long-wave
crests (troughs), as they are in the prior steady solutions
\citep{longuet1960changes,longuet1987propagation,zhang1990evolution}.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
\caption{
  Change in the short-wave action (left), wavenumber (middle), and steepness (right)
  relative to their initial values as function of initial long-wave phase and time.
  $k_L = 1$, $k = 10$, $\varepsilon_L = 0.1$.
}
\label{fig:modulation_3panel_infinite}
\end{figure}

Why does this unsteady growth of short-wave action occur?
\citet{peureux2021unsteady} explain that the change in advection velocity due to
the modulation of the short-wave group speed yields an additional amplification
of the short-wave action.
That is, the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ is
responsible for the instability.
However, the instability only occurs if the short waves are initialized as a
uniform field.
If they are initialized instead as a periodic function of the long-wave phase
such that their wavenumber (and optionally, action) is higher on the crests
and lower in the troughs, akin to the prior steady solutions, the instability
vanishes and the short-wave modulation pattern locks-in to the long-wave
crests.

@@

Why is it, then, that the solution for short-wave modulation is not only
quantitatively, but also qualitatively, different depending on the short-wave
initial conditions?
It may at first seem that there is something special about the uniform initial
conditions that makes its solutions grow unsteadily and indefinitely.
\citet{peureux2021unsteady} described this scenario as "...the sudden appearance
of a long wave perturbation in the middle of a homogeneous short wave field".
However, such a scenario is not realistic in the ocean where waves tend to
organize into groups due to dispersion.
To test whether the sudden onset of long-wave forcing causes the unstable growth,
we repeat the numerical simulation of Fig.
\ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
(and consequently, the other long-wave properties including the orbital
velocities and the local steepness) for the first five periods:
$a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
The key distinction that arises with the introduction of the ramp is that the
modulation pattern of short waves is now locked-in to the long-wave crests,
and their indefinite steepening vanishes.
The stabilization effect of the long-wave ramp is more readily assessed in
Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
steepness modulation as function of time over 30 long-wave periods, for the
infinite long-wave train case and the case of applying a linear ramp.
Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
inhomogeneity ($N \frac{\partial C_g}{\partial x}$) tendencies reveals that
after only one long-wave period, the propagation and inhomogeneity tendencies
are unbalanced and amplify the steepening of short waves
(Fig. \ref{fig:inhomogeneity_tendencies}).
\citet{peureux2021unsteady} correctly pointed out that the unstable
growth of short-wave steepness occurs due to the resonance of the inhomogeneity
tendency with the short-wave action perturbation.
However, in the case of a gentle ramp-up of the long-wave forcing, the
propagation and inhomogeneity tendencies mostly balance each other and allow
the short-wave steepening to lock-in to the long-wave crests.

@@

The solutions presented here, analytical and numerical alike, are valid only
for certain scales of short and long waves.
Specifically, the wave action conservation equation, introduced by
\citet{bretherton1968wavetrains}, requires that the ambient velocity is slowly
varying on the time scales of short waves, \textit{i.e.}
$|\partial U / \partial x| \ll k |U|$
and $|\partial U / \partial t| \ll \sigma |U|$.
This implies $k_L \ll k$ and $|\sigma_L| \ll \sigma|$, respectively.
Another requirement is for the short-waves to remain linear, which implies
$a k \ll 1$.
Furthermore, the short waves examined here do not include the effects of
surface tension (gravity-capillary waves) or intermediate or shallow water
(although the solutions can be extended to those conditions, as
\citet{phillips1981dispersion} did, for example).
Within these limits, the crest-action conservation equations provide an
elegant and relatively simple framework to study the hydrodynamic modulation
of short waves by longer waves.
This approach, of course, isolates this particular process from other
dominant modulation and nonlinear transfer processes, such as wave growth
by wind, wave dissipation, and aerodynamic sheltering by the longer waves.

@@

Despite several studies that provided qualitatively similar steady solutions
for the short-wave modulation by long waves
\citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution},
it was not obvious from the governing equations that the fully-nonlinear
steady solutions should exist.
\citet{peureux2021unsteady} examined this problem numerically and found that
the solutions are stable only if the short-wave field is initialized from
a prior steady solution of the form of that from \citet{longuet1960changes}.
If the short-wave field is instead initialized from a uniform field, they
found that the short-wave amplitude, and thus steepness as well, grow
unsteadily.
However, here we demonstrate that the unsteady steepening of short
waves only occurs in the case of a sudden appearance of long waves, where
the velocity fields induced by the long waves act in their full capacity
on the short waves that have not yet had time to lock-in to the long-wave
crests.
In such a hypothetical scenario, sudden forcing on short waves causes the
modulation pattern to detach itself from the long-wave crests and to propagate
at the short-wave group speed.
When this occurs, the modulated short-waves are no longer in a stabilizing
resonance with the surface convergence and divergence induced by the long-wave
orbital velocity, and each long-wave passage further destabilizes the short-wave
field, causing it to steepen indefinitely.

Another important consideration of the hydrodynamic modulation process is
whether it should somehow be accounted for in phase-averaged spectral wave
models \citep{group1988wam,tolman1991third,donelan2012modeling} used for global
and regional wave forecasting.
Currently, most spectral wave models used for research or forecasting
do not consider the preferential steepening of short waves in the context
of their dispersion properties or the wind input
(\textit{i.e.} via the short-wave phase speed).
Some dissipation source functions use the cumulative mean squared slope
to quantify preferential dissipation of short waves due to this effect
\citep{donelan2012modeling,romero2019distribution}.
In the early days of spectral wave modeling, \citet{willebrand1975energy}
examined the effect of a wave spectrum on the short-wave group velocity,
and found the effect of up to several percent modulation.
For a practical implementation, this would incur an additional loop over all
frequencies longer than the considered short-wave frequency.
It was thus not deemed practical to include this effect in wave models at the time.
However, with the recent advances in computing power, machine learning for
acceleration of complex functions, and phase-resolved modeling of short waves,
considering the effect of wave-modulation beyond the simple cumulative mean
squared slope deserves re-examination.

@@

\backsection[Funding]{
  Milan Curcic was partially supported by the National Science Foundation
  under grant number AGS-1745384.
}

>>>>

All modulation solutions mentioned thus far are steady in the reference frame
of the long wave---short waves become shorter and higher (and thus, steeper)
preferentially on the long-wave crests.
Conversely, they elongate and flatten preferentially in the long-wave troughs.
Recently, \citet{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are appropriate by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario in
which a long-wave train suddenly appears to modulate a uniform field of short
waves.
The numerical simulations stabilise if the short-wave field is initialised
from existing steady solutions, like those by \citet{longuet1960changes}.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.
Looking at the conservation equations alone, it is not obvious that it is.

In this paper, we revisit this problem from linear wave theory but consider
the long-wave slope to be significant enough to require evaluating the
short-waves at the long-wave surface rather than at the mean water level.
This allows us to derive alternative, steady, non-linear solutions for the
modulation of short waves.
The analytical solutions yield modulations that are stronger than the original
solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
long-wave steepness up to $\approx 0.2$.
To validate the steadiness of the analytical solutions, we perform numerical
simulations of the full wave crest and action conservation laws, and compare
the results to the analytical solutions presented here, as well as the prior
numerical results.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarised in Table
\ref{table:modulation_literature_summary}.
We first present the governing equations in \S\ref{section:governing_equations},
and then the analytical solutions for the modulation of short waves by long waves
in \S\ref{section:analytical_solutions}.
Then, we describe the numerical simulations and their results in
\S\ref{section:numerical_solutions}.
Finally, we discuss the results and their implications in
\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.

@@

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L \cos{\psi}$,
where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
angular frequency.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearising the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta$ and the
linearised conservation of wave crests in (\ref{eq:wavenumber}) results in
excess short-wave wavenumber being created.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.

@@

Let us now examine the contributions of the different terms to the short-wave
effective gravity modulation.
Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
a variety of the gravity modulation forms.
This value of $\varepsilon_L$ allows easier visualisation of the modulation
differences, while still being relatively moderate.
To the first order in $\varepsilon_L$, the gravity modulation is out of phase
with the long-wave elevation, with the short waves experiencing a lower effective
gravity on the long-wave crests and a higher effective gravity in the troughs.
Evaluating the orbital accelerations at $z=\eta_L$ causes a gravity reduction
around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
the exponential change in the orbital accelerations away from the mean water level
(Fig. \ref{fig:effective_gravities}b).
The difference between the linear and Stokes long waves at this steepness is
negligible.
The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
Naturally, there is no curvature exactly at the crests and the troughs, and its
effects are largest at the front and rear faces of the long wave, where the
covariance between the local slope and the horizontal accelerations are largest.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/fig_effective_gravities.png}
\caption{
  (a) Analytical solutions for the effective gravitational acceleration
  modulation by long waves, as function of long-wave phase, for $\varepsilon_L = 0.2$.
  Black is for linear wave evaluated at $z=0$,
  blue and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and without curvature effects,
  and orange and green are for linear and Stokes waves, respectively, evaluated at $z=\eta$ and with curvature effects.
  (b) As in (a) but normalised by the modulation of the linear wave evaluated at $z=0$, $g_1$.
}
\label{fig:effective_gravities}
\end{figure}

@@

The modulation of short-wave amplitude can be derived in a similar way as we did
for the wavenumber, except that here we linearise the wave action balance
(\ref{eq:wave_action2}) and integrate it in time:

@@

To quantify the contribution of the nonlinear terms in (\ref{eq:wavenumber})
and (\ref{eq:wave_action2}), and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are discretised using
second-order central finite difference in space and integrated in time using the
fourth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, thus neglecting the Stokes drift
induced by long waves on short-wave groups
\citep{stokes1847,van2018stokes,monismith2020stokes}.
Their gravity modulation also does not consider the nonlinear or the curvature
effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the more sophisticated MUSCL4 scheme \citep{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The maximum relative error of the centred finite difference in this model
is $\approx 0.066\%$.
We show in the next section that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

The numerical equations here are integrated in a fixed reference frame,
rather than that moving with the long-wave phase speed.
Either approach produces equivalent modulation results, however a fixed
reference frame may allow for a more intuitive interpretation of the results.
This difference is important to keep in mind when comparing the results
of \citet{peureux2021unsteady} and those presented here; in their Fig. 1 the
long-wave phase is fixed and the short waves are moving leftward
with the speed of $C_{pL} - C_g - U$ (neglecting group speed inhomogeneity),
whereas in the figures that follow, the long wave is moving rightward with its
phase speed $C_{pL}$ and the short waves (that is, their action) are moving
rightward with the speed of $C_g + U$ (neglecting group speed inhomogeneity).
The crest and action conservation equations are integrated in the curvilinear
coordinate system in which the horizontal coordinate follows the long-wave
surface, akin to \citet{zhang1990evolution}.

@@

The first set of simulations is performed for the long-wave steepness of
$\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
The long-waves are initialised from rest ($a_L = 0$) and gradually ramped up
to their target steepness over 5 long-wave periods, thus emulating the passage
of a long-wave group
(the importance of which we discuss in more detail in the next subsection).
At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
and steepness modulation are all remarkably similar to the analytical solutions
(Fig. \ref{fig:numerical_solutions}a,d,g).
This suggests that at low long-wave steepness that the steady analytical solution
is a reasonable proxy for the fully-nonlinear crest-action conservation
equation solutions.
The numerical solutions diverge more notably from the analytical ones at
moderate steepness of 0.2 (Fig. \ref{fig:numerical_solutions}b,e,h), and are
considerably different at very high steepness of 0.4
(Fig. \ref{fig:numerical_solutions}c,f,i).
These differences suggest that at high steepness the inhomogeneity of
gravitational acceleration and group speed, otherwise ignored in the steady
solutions, become important.

@@

The key result of \citet{peureux2021unsteady} is that of unsteady steepening of
short waves when a homogeneous field of short waves is forced by an infinite
long-wave train.
The steepening is driven by the progressive increase in the action of short
waves (as opposed to their wavenumber), and is centred on the short-wave group
that begins its journey in the trough of the long wave.
This is because the unmodulated short waves that begin their journey in the
trough first enter the convergence region before reaching the divergence region.
Inversely, the short waves that begin around the crest of the long-wave first
enter the divergence region and will thus have their action and wavenumber
decreased.
We reproduce this behaviour in Fig. \ref{fig:modulation_3panel_infinite}, which
shows the evolution of the short-wave action, wavenumber, and steepness
modulation (here defined as the change of a quantity relative to its initial value).
The long waves have $\varepsilon_L = 0.1$ and $k_L = 1$, and the short waves
have $k = 10$.
After 10 long-wave periods, the short-wave action is approximately doubled,
consistent with \citet{peureux2021unsteady}.
Another important property of this solution is that the short-wave group that
experiences peak amplification (dampening) are not locked-in to the long-wave
crests (troughs), as they are in the prior steady solutions
\citep{longuet1960changes,longuet1987propagation,zhang1990evolution}.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_modulation_3panel_infinite_wave_train.png}
\caption{
  Change in the short-wave action (left), wavenumber (middle), and steepness (right)
  relative to their initial values as function of initial long-wave phase and time.
  $k_L = 1$, $k = 10$, $\varepsilon_L = 0.1$.
  The black contour follows the value of 1, which corresponds to no change
  from the initial values.
}
\label{fig:modulation_3panel_infinite}
\end{figure}

Why does this unsteady growth of short-wave action occur?
\citet{peureux2021unsteady} explain that the change in advection velocity due to
the modulation of the short-wave group speed yields an additional amplification
of the short-wave action.
That is, the inhomogeneity term $N \dfrac{\partial C_g}{\partial x}$ is
responsible for the instability.
However, the instability only occurs if the short waves are initialised as a
uniform field.
If they are initialised instead as a periodic function of the long-wave phase
such that their wavenumber (and optionally, action) is higher on the crests
and lower in the troughs, akin to the prior steady solutions, the instability
vanishes and the short-wave modulation pattern locks-in to the long-wave
crests.

@@

Why is it, then, that the solution for short-wave modulation is not only
quantitatively, but also qualitatively, different depending on the short-wave
initial conditions?
It may at first seem that there is something special about the uniform initial
conditions that makes its solutions grow unsteadily and indefinitely.
\citet{peureux2021unsteady} described this scenario as "...the sudden appearance
of a long wave perturbation in the middle of a homogeneous short wave field".
However, such a scenario is not realistic in the ocean where waves tend to
organise into groups due to dispersion.
To test whether the sudden onset of long-wave forcing causes the unstable growth,
we repeat the numerical simulation of Fig.
\ref{fig:modulation_3panel_infinite}, except for a linear ramp applied to $a_L$
(and consequently, the other long-wave properties including the orbital
velocities and the local steepness) for the first five periods:
$a_L(t) = min(a_L, a_L t / \left(5 T_L\right))$, where $T_L$ is the long-wave period.
This result is shown in Fig. \ref{fig:modulation_3panel_ramp}.
The key distinction that arises with the introduction of the ramp is that the
modulation pattern of short waves is now locked-in to the long-wave crests,
and their indefinite steepening vanishes.
The stabilisation effect of the long-wave ramp is more readily assessed in
Fig. \ref{fig:unsteady_growth_timeseries}, which shows the maximum short-wave
steepness modulation as function of time over 30 long-wave periods, for the
infinite long-wave train case and the case of applying a linear ramp.
Comparing the wave action propagation ($C_g \frac{\partial N}{\partial x}$) and
inhomogeneity ($N \frac{\partial C_g}{\partial x}$) tendencies reveals that
after only one long-wave period, the propagation and inhomogeneity tendencies
are unbalanced and amplify the steepening of short waves
(Fig. \ref{fig:inhomogeneity_tendencies}).
\citet{peureux2021unsteady} correctly pointed out that the unstable
growth of short-wave steepness occurs due to the resonance of the inhomogeneity
tendency with the short-wave action perturbation.
However, in the case of a gentle ramp-up of the long-wave forcing, the
propagation and inhomogeneity tendencies mostly balance each other and allow
the short-wave steepening to lock-in to the long-wave crests.

@@

The solutions presented here, analytical and numerical alike, are valid only
for certain scales of short and long waves.
Specifically, the wave action conservation equation, introduced by
\citet{bretherton1968wavetrains}, requires that the ambient velocity is slowly
varying on the time scales of short waves, \textit{i.e.}
$|\partial U / \partial x| \ll k |U|$
and $|\partial U / \partial t| \ll \sigma |U|$.
This implies $k_L \ll k$ and $|\sigma_L| \ll \sigma|$, respectively.
Another requirement is for the short-waves to remain linear, which implies
$a k \ll 1$.
Furthermore, the short waves examined here do not include the effects of
surface tension (gravity-capillary waves) or intermediate or shallow water
(although the solutions can be extended to those conditions, as
\citet{phillips1981dispersion} did, for example).
Within these limits, the crest-action conservation equations provide an
elegant and relatively simple framework to study the hydrodynamic modulation
of short waves by longer waves.
This approach, of course, isolates this particular process from other
dominant modulation and nonlinear transfer processes, such as wave growth
by wind, wave dissipation, and aerodynamic sheltering by the longer waves,
that are otherwise present in most measurements
\citep{plant1986two,laxague2017gravity}.

@@

Despite several studies that provided qualitatively similar steady solutions
for the short-wave modulation by long waves
\citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution},
it was not obvious from the governing equations that the fully-nonlinear
steady solutions should exist.
\citet{peureux2021unsteady} examined this problem numerically and found that
the solutions are stable only if the short-wave field is initialised from
a prior steady solution of the form of that from \citet{longuet1960changes}.
If the short-wave field is instead initialised from a uniform field, they
found that the short-wave amplitude, and thus steepness as well, grow
unsteadily.
However, here we demonstrate that the unsteady steepening of short
waves only occurs in the case of a sudden appearance of long waves, where
the velocity fields induced by the long waves act in their full capacity
on the short waves that have not yet had time to lock-in to the long-wave
crests.
In such a hypothetical scenario, sudden forcing on short waves causes the
modulation pattern to detach itself from the long-wave crests and to propagate
at the short-wave group speed.
When this occurs, the modulated short-waves are no longer in a stabilising
resonance with the surface convergence and divergence induced by the long-wave
orbital velocity, and each long-wave passage further destabilises the short-wave
field, causing it to steepen indefinitely.

Another important consideration of the hydrodynamic modulation process is
whether it should somehow be accounted for in phase-averaged spectral wave
models \citep{group1988wam,tolman1991third,donelan2012modeling} used for global
and regional wave forecasting.
Currently, most spectral wave models used for research or forecasting
do not consider the preferential steepening of short waves in the context
of their dispersion properties or the wind input
(\textit{i.e.} via the short-wave phase speed).
Some dissipation source functions use the cumulative mean squared slope
to quantify preferential dissipation of short waves due to this effect
\citep{donelan2012modeling,romero2019distribution}.
In the early days of spectral wave modelling, \citet{willebrand1975energy}
examined the effect of a wave spectrum on the short-wave group velocity,
and found the effect of up to several percent modulation.
For a practical implementation, this would incur an additional loop over all
frequencies longer than the considered short-wave frequency.
It was thus not deemed practical to include this effect in wave models at the time.
However, with the recent advances in computing power, machine learning for
acceleration of complex functions, and phase-resolved modelling of short waves,
considering the effect of wave-modulation beyond the simple cumulative mean
squared slope deserves re-examination.

@@

\backsection[Funding]{
  Milan Curcic was partially supported by the National Science Foundation
  under the grant number AGS-1745384.
}
